---
title: "HW6"
output: md_document
date: "2024-02-07"
---
# Birthday Question

## 1

The probability is 0.9875. 

## 2 

Set seed is used in this code to control the random number generation.

## 3

R represents the number of times the random class will be generated. For example, if R is 10 then 10 classes of birthdays will be generated and looked through for duplicates.

## 4

This solutions assumes that each day is equally likley for a birthday. Therefore, the code is ignoring the fact that less people give birth on Friday the 13th, it all together ignores leap year, and overall ignores human factors.  
## 5

They both gave very similar answers fro a class size of 28. the lines on the graph match up very well to the probability from version 1. 

## 6

See graph below

## 7

I think that first_duplicate is returning the first number that produces a duplicate. for example, for 28 the code spit out 11, which means 11 classes had to be randomly produced before a duplicate was found. 

```{r}
generate_class <- function(class_size){
  sample(1:365, class_size, replace=TRUE)
}

birthday_roster <- generate_class(28)

check_birthday <- function(birthday_roster){
  birthday_roster |> duplicated() |> any()
  #use sum() or any() command to count up
}

set.seed(230583)
R <- 10000
replicates <- replicate(R, generate_class(28) |> check_birthday())
replicates
mean(replicates)
```

```{r}

R <- 10000

first_duplicate <- function(){
    sample(1:365, 366, replace=TRUE) |>
    duplicated() |>
    which() |>
    min()
}

fd1 <- replicate(R, first_duplicate())
plot(ecdf(fd1), main = "Probability of shared birthday", xlab = "Class size")

c28 <- 0.65
abline(h=c28, v=28)
```
