---
title: "HW 2"
output: md_document
date: "2024-01-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

> If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6, and 9. The sum of these multiples is 23.Find the sum of all the multiples of 3 or 5 below 1000.

**First, count a set from 3-1000 by 3s. Then count a set 5-1000 by 5s. Then sum the outputs of both sets.** 

```{r}
a <- 1:999
a[(a %% 3 == 0) | (a %% 5 == 0)]
```

## Problem 2

> Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first terms 10 will be: 1, 2, 3, 5, 8, 13, 21, 34, 55, 89. By considering the terms in the Fibonacci sequence whose values do not exceed four million, find the sum of the even-valued terms.

**Loop through each output to continuously add the last two digits together. Have the function break when 4 million is reached**

``{r}``


## Problem 3

> problem

**loop through 600851475143 by dividing each following output by 2. If the output is 0, continue to divide again by two. If not 0, divide by 3; if that output is 0, continue. If not 0, divide by 5. Continue the process stated with each prime number until a final number is reached**

``{r}``

