---
title: "hw3"
output: html_document
date: "2024-01-31"
---

# Question 1

## What is a working directory?

a working directory contains all data, rmd scripts, and generated outputs. Create a folder for each homework assignment and put all relavent files into that folder. For example, for this assignment I have created a folder called "HW3". Once I open the folder, all the files I create will be saved into HW3. I can also use the R studio feature session -> Set working directory -> To source file location. 

## What are the best practices for setting and using the working directory? 

 We work in teams and share code a lot of the time in data science. An absolute file path is something like (c:/Users/Rxb8yt/project/HW3/data/survey.csv). when you write this into the code script, it will not exist in this exact location on another persons device. Therefore, do not use exact file path locations. Instead, use a relative file path (../data/survey/csv) which will create a file path relative to wear the working directory is.(.. means go up a level) 

# Q2

```{r}
name <- "Olivia Park"
```

#Q3

```{r}
ls()
```
# Q4

```{r}
v1 <- c(FALSE, TRUE, FALSE, TRUE, TRUE)
v2 <- c(1, 2, 3, 4, 5)
v3 <- c(1.1, 2.2, 3.3, 4.4, 5.5)
v4 <- c("a", "b", "c", "d", "e")
```

#Q5

```{r}
V <- list(v1, v2, v3, v4)
V
```

#Q6

```{r}
str(V)
```

#Q7

```{r}
matrix(data = 1:20, nrow = 4, ncol = 5)
```
#Q8

```{r}
seq(22, 44)
```

#Q9

```{r}
a <- cbind(1 ,1:10)
a
```

#Q10

```{r}
v3[3]
```

#Q11

```{r}
d1 <- read.csv("http://hbiostat.org/data/repo/2.20.Framingham.csv")
d1[ ,4]
```

#Q12

```{r}
d1[5, ]

```

#Q13

```{r}
d1["bmi"]
```

#Q14
The pipe operator pushes the change made in one line forward into the next line so that an aspect of the data frame is altered at once.
```{r}
require(dplyr)

# Old way
sum(log(sqrt(select(d1,age)),base=10))

# OG magrittr pipe
d1 %>% 
  pull(age)  %>% 
  sqrt  %>% 
  log(base=10)  %>% 
  sum

d1 %>% 
  lm(sbp ~ dbp, data = .)  # Use the dot if the piped objects needs to go to an input other than the first

d1 %>% 
  split(.$sex) %>%  # Can use the input multiple times
  lapply(function(x){x$age %>% mean})

# New base R pipe
d1 |> 
  pull(age) |> 
  sqrt() |> # Note the parentheses have to be used
  log(base=10) |> 
  sum()

d1 |>  
  lm(sbp ~ dbp, data = _) # Use `_` instead of `.` with new pipe

```
#Q15

```{r}

prime <- function(x){
  ps <- primes::generate_primes(min = 2, max = ceiling(sqrt(x)))
  for(i in ps) if(x %% i == 0) return(FALSE)
  TRUE
}

k <- 0
for(i in 0:9999){
    phone_number <- 4349240000 + i
    if(prime(phone_number)) {
      k <- k + 1
    }
}
k
```

#Q16

```{r}
prime_phonenumbers <- c()
for(i in 0:9999){
    phone_number <- 4349240000 + i
    if(prime(phone_number)) {
    prime_phonenumbers <- c(prime_phonenumbers, phone_number)
    }
     }
prime_phonenumbers
```