---
title: "Homework 9"
output: md_document
date: "2024-02-26"
---

# 1

Add up cell probabilties
```{r}
one <- .08 + .18 + .11 + .19 
one
```

# 2

Calculate the conditional row probability 
```{r}
two <- .11 /( .11 + .19) 
two
```

# 3

Add up all cells where statement is true
```{r}
three <- .03 +.11 +.18 + .11 + .19 +.17 
three
```
# 4 

conditional column probability
```{r}
pa <- 0.03+0.08+0.11+0.13 #margin
(0.03+0.08)/pa
```
# 5
```{r}
Probability <- 1 - (365/365 * 364/365 * 363/365 * 362/365 * 361/365 * 360/365 * 359/365 * 358/365 * 357/365 * 356/365 * 355/365)
Probability
```
```{r}
generate_class <- function(class_size){
  sample(1:365, class_size, replace=TRUE)
}

birthday_roster <- generate_class(11)

check_birthday <- function(birthday_roster){
  birthday_roster |> duplicated() |> any()
  #use sum() or any() command to count up
}

set.seed(230583)
R <- 10000
replicates <- replicate(R, generate_class(11) |> check_birthday())
replicates
mean(replicates)
```
> The probabilities are pretty close but not exact. The forumla outputs .1411 and the simulation outputs .1492

# factorial function - wont work with formula because numbers are too big - use logs and exponents
```{r}
k <- 11
1 - exp(lfactorial(365) - lfactorial(365 - k) - k*log(365))
```


# when running this code, we can see that the anaylitical soultion and code matches up perfectly

```{r}
R <- 10000

first_duplicate <- function(){
    sample(1:365, 366, replace=TRUE) |>
    duplicated() |>
    which() |>
    min()
}

fd1 <- replicate(R, first_duplicate())
plot(ecdf(fd1), main = "Probability of shared birthday", xlab = "Class size")

k <- 11
a <- 1 - exp(lfactorial(365) - lfactorial(365 - k) - k*log(365))
abline(h=a, v=11)

bp <- function(k) 1 - exp(lfactorial(365) - lfactorial(365 - k) - k*log(365))
cs1 <- 1:80
bp1 <- bp(cs1)
lines(cs1, bp1, col = "red", lwd = 4)

ecdf(fd1)(cs1) - bp1

```