---
title: "HW5"
output: md_document
date: "2024-02-06"
---

```{r}
big_well <- function(N) 100*rbinom(N, 1, 0.1)
sma_well <- function(N)  10*rbinom(N, 10, 0.1)

set.seed(23459874) # What does set.seed do?
big_well(10)

##  [1] 100 100   0   0   0   0   0   0   0   0

sma_well(10)

##  [1] 10  0  0 10  0  0  0  0 20  0

big_well(1E6) |> 
  table() 

## 
##      0    100 
## 899848 100152

big_well(1E6) |>
  table() |>
  proportions()

## 
##       0     100 
## 0.90008 0.09992
```

# 1: Write code that generates the long-run proportions for the small well investment opportunity.

```{r}
sma_well(1E6) |>
  table() |>
  proportions()
```

# 2: Create a figure which visualizes the long-run proportions

```{r}
# new package needed 
# install.packages("ggplot2")
library(ggplot2)

# place proportions into a new variable 

sma_proportions <- sma_well(1E6) |>
  table() |>
  proportions()

# create a data frame to input into a table

sma_df <- data.frame(value = as.numeric(names(sma_proportions)),
                     proportion = as.numeric(sma_proportions))
# use ggplot to graph

ggplot(sma_df, aes(x = factor(value), y = proportion)) +
  geom_bar(stat = "identity") +
  labs(title = "Long-run proportion of the small well investment opportunity",
       x = "Value",
       y = "Proportion")

```

# Questions on Chapter 3

## 1 
K represents knowledge. An example of how K can change probabilty how be initally thinking I will go to class, and this would change once I knew it was going to rain, because my chances of wanting to go to class would decrease.

## 2 
Ec represents a set of outcomes that are not E. Ec is all other outcomes. Another way to think about Ec would be as a compliment to event E. 

## 3 
The P(E is odd) would be 0.4, because there are only two options, odd or even.

